How cryptic is cryptic diversity? Machine learning approaches to classifying morphological variation in the Pacific Pond Turtle (\textit{Emys marmorata})

### 2016-03-03 doc/resubmit_letter.tex
ken comments. editor and first reviewer almost finished. second reviewer hurts me inside.

@@

@@

  \textbf{My second criticism has to do with the choice of data set for the author’s ``clear-cut example''. In one view, this study doesn’t need a ``clear-cut'' example for comparison at all. This is the case if the question of the study is: ``Is there evidence for morphological differentiation in the plastra of E. marmorata, consistent with some previously-proposed classifications?'' But on another view, an appropriate comparator is necessary. This is the case if the question of the study is: ``Is there as much evidence in the plastra of E. marmorata for grouping members of this species into multiple subspecies, as there is in other turtle species which have already been grouped into multiple subspecies?''. And of course, a comparator is also necessary to answer the question: ``Can the methods used in this paper ever produce a taxonomically-appropriate clustering?''\\
    The authors clearly intend the latter for their study. They almost could not have chosen a set of taxa better-suited to successfully finding differences (maybe not at all within a given subfamily of turtles). The species they have chosen differ in pretty much all of the major factors that are known to affect the plastra of turtles: size, plastral kinesis, habitat, and degree of phylogenetic relatedness. Perhaps this is why they chose the taxa that they did, but if so, then I’m not sure that the results are particularly germane to the rest of the analyses. Of course the methods will produce an appropriate grouping for that set of taxa - I cannot imagine any published methods not doing so. So these results don’t really provide a useful comparator with the E. marmorata groupings; they just show that the methods used aren’t utterly abysmal. \\
    Instead, I think that the authors need to be able to answer the second question that I proposed above. They need to be able to show that these methods can successfully group turtle plastra which show the same magnitude of variation that is expected in subspecies or recently-diverged species or cryptic species in related groups. Grouping species within Emydinae won’t provide this. I’m not sure what the best comparisons would be. Perhaps Chrysemys subspecies, or C. picta versus C. dorsalis. Perhaps Terrapene species or subspecies. But the authors should pick and justify an appropriate comparison. And if the methods cannot successfully group these taxa either, then the relevance of their failure to group E. marmorata will be put in its proper context. \\
  The paper is otherwise strong. Again, I think the authors did an excellent job of synthesizing results from multiple methods, and the study is important (and, needless to say, appropriate for Systematic Biology). But I think the authors need to clean up their presentation and choose a more appropriate comparison or validation of the relevance of their results. }

>>>>

  The results of PAM clustering have been clarified. The biggest detail is remembering the confidence intervals around the mean. In the previous draft, only a single standard deviation was plotted which may lead to confusion. We've changed this so that two standard deviations are plotted as confidence intervals; this makes the result that no single classification scheme is optimal much more obvious.

@@

  These typographic errors and content omissions were fixed where applicable. Additionally, the sources of the schemes have been identified early in the Methods section.

@@

  354 points are plotted it is just that many of them are plotted on top of each other. We've made these points slightly transparent to help understand how many points there are and where they occur.

  Additionally, we've tested if the two-cluster PAM solution corresponded to the sexes of the specimens. While we did not have complete sex information for every specimen, we found that this clustering solution has no correspondence with sex.

  \textbf{My second criticism has to do with the choice of data set for the author’s ``clear-cut example''. In one view, this study doesn’t need a ``clear-cut'' example for comparison at all. This is the case if the question of the study is: ``Is there evidence for morphological differentiation in the plastra of E. marmorata, consistent with some previously-proposed classifications?'' But on another view, an appropriate comparator is necessary. This is the case if the question of the study is: ``Is there as much evidence in the plastra of E. marmorata for grouping members of this species into multiple subspecies, as there is in other turtle species which have already been grouped into multiple subspecies?''. And of course, a comparator is also necessary to answer the question: ``Can the methods used in this paper ever produce a taxonomically-appropriate clustering?''\\\\
    The authors clearly intend the latter for their study. They almost could not have chosen a set of taxa better-suited to successfully finding differences (maybe not at all within a given subfamily of turtles). The species they have chosen differ in pretty much all of the major factors that are known to affect the plastra of turtles: size, plastral kinesis, habitat, and degree of phylogenetic relatedness. Perhaps this is why they chose the taxa that they did, but if so, then I’m not sure that the results are particularly germane to the rest of the analyses. Of course the methods will produce an appropriate grouping for that set of taxa - I cannot imagine any published methods not doing so. So these results don’t really provide a useful comparator with the E. marmorata groupings; they just show that the methods used aren’t utterly abysmal. \\\\
    Instead, I think that the authors need to be able to answer the second question that I proposed above. They need to be able to show that these methods can successfully group turtle plastra which show the same magnitude of variation that is expected in subspecies or recently-diverged species or cryptic species in related groups. Grouping species within Emydinae won’t provide this. I’m not sure what the best comparisons would be. Perhaps Chrysemys subspecies, or C. picta versus C. dorsalis. Perhaps Terrapene species or subspecies. But the authors should pick and justify an appropriate comparison. And if the methods cannot successfully group these taxa either, then the relevance of their failure to group E. marmorata will be put in its proper context. \\\\
  The paper is otherwise strong. Again, I think the authors did an excellent job of synthesizing results from multiple methods, and the study is important (and, needless to say, appropriate for Systematic Biology). But I think the authors need to clean up their presentation and choose a more appropriate comparison or validation of the relevance of their results. }

  You are correct that our goal was the to address the second question. Following your suggestion, we've included an additional dataset looking at plastron shape variation between the two subspecies \textit{Trachemys scripta}. Between these two additional datasets we believe we have demonstrated that our approach can work when classes are actually distinct.
